A Note for Solving String Constraints with Split Function

### 2023-10-23 2023-9-4_note/extend_split/note.tex
daily update the note of string array

\newcommand*{\aut}{\mathcal{A}}
\newcommand*{\Int}{\mathbb{Z}}
\newcommand*{\myvec}[1]{\vec{#1}}
\newcommand{\cefaout}{\mathcal{O}}
\newcommand{\Lang}{\mathcal{L}}
\newcommand{\arrconcat}{\sharp}
\newcommand{\denghang}[1]{{\color{teal}\textbf{DH:} #1 \textbf{:HD}\color{black}}}
\newcommand{\comment}[1]{{\color{blue}\textbf{MARK:} #1 \textbf{:KRAM}\color{black}}}
\newcommand{\todo}[1]{{\color{orange}\textbf{TODO:} #1 \textbf{:ODOT}\color{black}}}
\newcommand{\mysplit}{\mathsf{split}}
\newcommand{\myjoin}{\mathsf{join}}
\newcommand{\mylen}{\mathsf{len}}
\newcommand{\myread}{\mathsf{read}}
\newcommand{\mywrite}{\mathsf{write}}
\newcommand{\myset}[1]{\{#1\}}

\theoremstyle{remark}
\newtheorem{definition}{Defination}[section]

@@

\subsection{Special CEFA for String Array}
%%
We define a special CEFA for string array. The CEFA uses a preserved character $``\arrconcat"$ as $array$ elements separator (note that $\arrconcat\not\in\Sigma$). The speciality of $\aut_s$ rely on  its construction.
%%
\paragraph{Array of fixed length and constant index} Considering an $array$ $A$ with fixed length $m$, we assume that each array element $A[i]$ $(i\in[1,m])$ is constrained by a CEFA $\aut_i\equiv (R_i, Q_i, \Sigma, \delta_i, I_i, F_i, \alpha_i)$. Note that the index of the array starts from 1.  Then we can construct the CEFA of the array $A$ as $\aut_s\equiv (R_s, Q_s, \Sigma\cup\{\arrconcat\}, \delta_s, I_s, F_s, \alpha_s)$, where:
\begin{itemize}
    % \item $R_s = \bigcup\limits_{1\leq i \leq m} R_i$,
    \item $R_s = \bigcup\limits_{1\leq i \leq m} R_i\cup \myset{r}$ for a fresh register r,
    \item $Q_s = \bigcup\limits_{1\leq i \leq m} Q_i\cup \myset{q_0, q_f}$ for fresh states $q_0$ and $q_f$,
    \item $\delta_s$ is composed of four transitions set
          \begin{itemize}
              \item  $\{(q_0, \arrconcat, q_1, (\myvec{0}, 1))\mid q_1\in I_1\}$,
              \item  $\myset{(q_m, \sharp, q_f, (\myvec{0}, 0))\mid q_m\in F_m}$,
              \item  $\{(q, \arrconcat, q', (\myvec{0}, 1)) \mid q\in F_i, q'\in I_{i+1}, i\in[1,m]\}$ where $ I_{m+1}=\emptyset$,
              \item  and $\{(q, a, q', (\myvec{0},\myvec{v_i},\myvec{0}))\mid (q, a, q', \myvec{v_i})\in \aut_i, i\in[1,m]\}$,
          \end{itemize}
          where $(\myvec{0}, 1)$ is a vector that increases the value of $r$ by 1 and remains other registers unchanged, and $(\myvec{0},\myvec{v_i},\myvec{0})$ is a vector that updates the value of $R_i$ by $\myvec{v_i}$ and remains other registers unchanged. Similarly for $(\myvec{0}, 0)$ and so on.
    \item $I_s = \myset{q_0}$, $F_s = \myset{q_f},$ and $\alpha_s = \bigwedge\limits_{1\leq i \leq m} \alpha_i$.
          % and $\alpha_s = \bigwedge\limits_{1\leq i \leq m} \alpha_i$.
\end{itemize}
The resulting CEFA is shown in fig \ref{fig:fixed_len_cons_idx}, where $q_i\in I_i$ and $q_i'\in F_i$ for each $i\in [1,m]$. The main idea of the construction above is using $\arrconcat$ to separate each element of the array and one register to record the index of each element. We introduce two fresh states $q_0$ and $q_f$ to ensure each element is started and ended with at least one $\arrconcat $, which is beneficial to the construction of pre-images of string constraints (see \ref*{sec:pre_image} for more details). The accepting condition $\alpha_s$ is the conjunction of accepting conditions of each element.
% It is obviously that the accepting word $w$ of $\aut_s$ is in language $\mathcal{L}(\aut_1)\arrconcat\mathcal{L}(\aut_2)\ ... \ \arrconcat \mathcal{L}(\aut_{m}) $.
\begin{figure}[H]
    \centering
    \usetikzlibrary {automata, positioning, fit}
    \begin{tikzpicture}[initial text =, initial distance=3ex,
            node distance=1.2cm, auto,
            state/.style={circle, draw, inner sep=0pt, minimum size=5mm},
            accepting by double]

@@

    \end{tikzpicture}
    \caption{The CEFA of $A$ with accepting condition $\alpha_s = \bigwedge\limits_{1\leq i \leq m} \alpha_i$}
    \label{fig:fixed_len_cons_idx}
\end{figure}
The construction of $\aut_s$ can be used to formalize the array containing constant string. For example, for a 2-length array $A$ constrained by $``a" = \myread(A, 1)\wedge ``b" = \myread(A, 2) \wedge \mylen(A) = 2$ , where $\myread(A, m)$ read the  element of $A$ at constant index $m$ and $\mylen(A)$ is the length of $A$, we can construct the automaton of $A$ as following: \\
$(\emptyset, \{q_0,q_1,q_1',q_2,q_2'\}, \{a,b,\arrconcat\}, \{q_0\xrightarrow[(1)]{\arrconcat}q_1, q_1\xrightarrow[(0)]{a}q_1', q_1'\xrightarrow[(1)]{\arrconcat}q_2, q_2\xrightarrow[(0)]{b}q_2'\}, \{q_0\}, \{q_2'\}, \top)$. 
%%
\paragraph{Array of varible length and index} \todo{TODO}

\section{Straight-Line String Theory with Integer and Array $SL_{I,A,S} $}\label{sec:SL}
We defined a theory $SL_{I,A,S}$ with three sorts: \textbf{string}, \textbf{integer} and \textbf{array}. The \textit{elements} of \textbf{array} can only be \textbf{string} sort, and the $index$ of \textbf{array} is \textbf{Integer} sort.
\subsection{Syntax}
The syntax of $SL_{I,A,S}$ is defined in table \ref{tab:syntax} where: $e$ is a regular expression; $\mylen(A)$ is the length of array $A$, $\mylen(x)$ is the length of string $x$; $\myread(A, m)$ read the  element of $A$ at index $m$, $\mywrite(B, x ,m)$ write string $x$ to array $B$ at index $m$;  $\myjoin(A,u)$ connect all of elements in array $A$ by connector $u$; $\mysplit(s,e)$ split string $x$ into ordered string array by separator $e$; $y\cdot z$ connects string $y$ and $z$; $\circ \in \{=,\not=, <, \leq, >, \geq \}$ is the set of arithmatic binary predicates.
%  For technical convenience, we assume that all assignments in F are \textit{Static Singly Assignment (SSA)} form.
\begin{table}
    \renewcommand{\arraystretch}{1.5}
    \begin{tabular}{l l l r}
        $x,y$ & $\cdots$ &                                                                       & String varibles          \\
        $m,n$ & $\cdots$ &                                                                       & Integer constants        \\
        $i,j$ & $\cdots$ &                                                                       & Integer varibles         \\
        $A,B$ & $\cdots$ &                                                                       & Array varibles           \\
        $t$   & $::=$    & $m \mid i \mid t+t \mid t-t \mid \mylen(A)\mid \mylen(x) $            & Integer term             \\
        $SC$  & $::=$    & $x=\myread(A, m) \mid A = \mywrite(B, x ,m) \mid x=\myjoin(A,u) \mid$ &                          \\
              &          & $A = \mysplit(x, e) \mid x= y\cdot z\mid x\in e$                      & String constraints       \\
        $F$   & $::=$    & $t\circ t \mid SC \mid F\wedge F $                                    & Formula of $SL_{I,A,S} $

    \end{tabular}
    \caption{Syntax of $SL_{I,A,S}$}
    \label{tab:syntax}
\end{table}
\subsection{Defination of Straight-Line Restriction }
We consider two sorts: \textbf{string} and \textbf{array}.
\begin{definition}[Relational constraints ]
    Relational constraints $\phi$ are defined by the following rules:
    \begin{align*}
        \phi \ ::= \  & x=y\cdot z \mid x=\myjoin(A,u) \mid A = \mysplit(x,e) \mid        \\
                      & x=\myread(A, m) \mid A = \mywrite(B, x ,m) \mid \phi \wedge \phi 
    \end{align*}
\end{definition}
\begin{definition}[Straight-line relational constraints]
    A relational constraint $\phi$ is straight-line if $\phi = \bigwedge\limits_{1\leq i \leq m} \chi_i = P_i$ such that
    \begin{itemize}
        \item $\chi_1,\cdots,\chi_m $ are mutally distinct string varibles and array varibles.
        \item For each $i\in [m] $, all the varibles in $P_i$ are in the set $\{\chi_1,\cdots,\chi_{m-1} \}$.
    \end{itemize}
\end{definition}
\begin{definition}[Straight-line formula]
    A formula $F$ of $SL_{I,A,S}$ is straight-line iff all relational constraints in $F$ are straight-line.
\end{definition}
% \begin{figure}[H]
%     % denote
%     % $u,v,\cdots$  \hfill{String constants}\\
%     $x,y,\cdots$  \hfill{String varibles}\\
%     $m,n,\cdots$  \hfill{Integer constants}\\
%     $i,j,\cdots$  \hfill{Integer varibles}\\
%     % $M,N,\cdots$  \hfill{Array constants}\\
%     $A,B,\cdots$  \hfill{Array varibles}\\

%     % term and expression
%     % $s \ ::= \ u \mid x$ \hfill{String term}\\
%     $t \ ::= \  m \mid i \mid t+t \mid t-t \mid |A|\mid |x| $ \hfill{Integer term}\\
%     % $A \ ::= \  M \mid X  $ \hfill{Array term}\\
%     % $\beta \ ::= \  m \mid |A| \mid |A|+m  $  \hfill{Bound expression}\\

%     % formula
%     % $LA \ ::= \ t\circ t \mid LA\wedge LA \mid \neg LA $  \hfill{Linear arithmatic}\\
%     $SC \ ::= \ x=read(A, m) \mid A = write(A, x ,m) \mid x=\myjoin(A,u)\mid A = \mysplit(x, e) \mid x\in e$   \flushright{String constraints}\\
%     % $AC \ ::= \ A = \mysplit(s,e) \mid A = A[\beta_1:\beta_2] \mid A = A[t\rightarrow s] \mid A\in \aut_s \mid A = \mathcal{T}(A) $ \flushright{Array constraints}\\ 
%     $F \ ::= \ t\circ t \mid SC \mid F\wedge F $ \hfill{Formula of $SL_{I,A,S} $}
%     \caption{Syntax of $SL_{I,A,S}$}
%     \label{fig:syntax}
% \end{figure}

% \section{Pre-image of String Constraints}
% \subsection{Array Assignment}
% Assume $A\in \aut_s$ and $\aut_s=(Q, \Sigma, \delta,q_0,F,R)$ construct like fig \ref{fig:special_nfa}.
% \begin{itemize}
%     \item $A = \mysplit(s,e)$: assume $e\in \mathcal{A}_e$, then we can construct $\mathcal{B}=(Q', \Sigma', \delta',q_0,F,R)$ like fig \ref{fig:pre_image_split} s.t $s\in B$.
%     \item $B = A[\beta_1:\beta_2]$:
% \end{itemize}

% % \begin{itemize}
% %     \item $e$ : A regular expression.
% %     \item $A[m]$ : Return the $element$ of $A$ at index $m$.
% %     % \item $A[t\rightarrow s]$ : Return a new array the same as $A$ except  that the $element$ of $A$ at index $t$ changes to string $s$.
% %     % \item $s\cdot s$ : String concatenation.
% %     \item $\myjoin(A,u) $ : Return a new string by concatenating all of $elements$ in array $A$, separated by a specified separator string $u$.
% %     \item $\mysplit(s,e) $ : Return a new array by dividing string $s$ into ordered substring array. The separator string is regular expression $e$.
% %     % \item $\mathcal{T}(A) $: Return a new array by applying transducer $\mathcal{T}$ on each $element$ of array $A$.
% %     % \item $A[\beta_1:\beta_2] $ Return a new array which is the subarray of $A$.
% % \end{itemize}
% % \subsection{Defination of Straight-Line Restriction }
% We consider two sorts: \textbf{string} and \textbf{array}.
% \begin{definition}[Relational constraints ]
%     Relational constraints $\phi$ are defined by the following rules:
%     \begin{align*}
%         \phi \ ::= \ x=s\cdot s \mid x=\myjoin(A,u) \mid A = \mysplit(s,e) \mid A = A[\beta_1:\beta_2] \mid A = A[t\rightarrow s] \mid \phi \wedge \phi
%     \end{align*}
% \end{definition}
% \begin{definition}[Straight-line relational constraints]
%     A relational constraint $\phi$ is straight-line if $\phi = \bigwedge\limits_{1\leq i \leq m} x_i = P_i$ such that
%     \begin{itemize}
%         \item $x_1,\cdots,x_m $ are mutally distinct.
%         \item For each $i\in [m] $, all the varibles in $P_i$ are either source varibles or varibles in set $\{x_1,\cdots,x_{m-1} \}$.
%     \end{itemize}
% \end{definition}
% \begin{definition}[Straight-line formula]
%     A formula $F$ of $SL_{I,A,S}$ is straight-line iff all relational constraints in $F$ are straight-line.
% \end{definition}

% \begin{figure}
%     \includegraphics[width=\linewidth]{pre_image_split.png}
%     \caption{pre-image of split operation}
%     \label{fig:pre_image_split}
% \end{figure}
% \bibliographystyle{abbrv} % 指定格式，
% % \bibliographystyle{apa} % 这是姓名年份应用格式
% \bibliography{ref} %  ../.bib表示bib的具体位置

\section{Pre-images of String Constraints}\label{sec:pre_image}
In this section, we define the pre-image of each string constraint formula.
\paragraph{- $i = \mylen(A)$:} The pre-image for this formula is a CEFA $\aut_A=(R_A, Q_A, \Sigma\cup\{\arrconcat\}, \delta_A, I_A, F_A, \alpha_A)$, where
\begin{itemize}
    \item $R_A = \{r\}$ for a fresh register $r$, $Q_A = \{q_0, q_f\}$, $\delta_A = \myset{q_0\xrightarrow[(0)]{\Sigma}q_0, q_0\xrightarrow[(1)]{\sharp}q_0, q_0\xrightarrow[(0)]{\arrconcat}q_f }$, $I_A = \myset{q_0}$, $F_A = \myset{q_f}$, and $\alpha_A \equiv r = i$.
\end{itemize}

@@

\paragraph{- $x = \myread(A, m)$:} Suppose that $x$ is constrained by $\aut_x=(R_x, Q_x, \Sigma, \delta_x, I_x, F_x, \alpha_x)$, then the pre-image for $A$ is a CEFA $\aut_A=(R_A, Q_A, \Sigma\cup\{\arrconcat\}, \delta_A, I_A, F_A, \alpha_A)$, where
\begin{itemize}
    \item $R_A = R_x$,
    \item $Q_A = Q_x\cup\myset{q_0, q_1, \cdots, q_{m}}$,
    \item $\delta_A$ is the union of six transitions set:
          \begin{itemize}
              \item $\myset{q_i\xrightarrow[(\myvec{0})]{\Sigma}q_i\mid i\in [1,m-1]}$,
              \item $\myset{q_i\xrightarrow[(\myvec{0})]{\arrconcat}q_{i+1}\mid i\in[0,m-2]}$,
              \item $\myset{q_{m-1}\xrightarrow[(\myvec{0})]{\arrconcat}q_0^x \mid q_0^x\in I_x}$,
              \item $\myset{q_f^x\xrightarrow[(\myvec{0})]{\arrconcat}q_m \mid q_f^x\in F_x}$,
              \item $\myset{q_m\xrightarrow[(\myvec{0})]{\arrconcat}q_m}$,
              \item $\myset{q\xrightarrow[(\myvec{v})]{a}q'\mid q\xrightarrow[\myvec{v}]{a}q'\in \delta_x}$.
          \end{itemize}
    \item $I_A = \myset{q_0}$, $F_A = \myset{q_m}$, and $\alpha_A = \alpha_x$.
\end{itemize}

@@

>>>>

\newcommand*{\aut}{\mathcal{A}}
\newcommand*{\Int}{\mathbb{Z}}
\newcommand*{\myvec}[1]{\vec{#1}}
\newcommand{\cefaout}{\mathcal{O}}
\newcommand{\Lang}{\mathcal{L}}
\newcommand{\arrconcat}{\sharp}
\newcommand{\denghang}[1]{{\color{teal}\textbf{DH:} #1 \textbf{:HD}\color{black}}}
\newcommand{\comment}[1]{{\color{blue}\textbf{MARK:} #1 \textbf{:KRAM}\color{black}}}
\newcommand{\todo}[1]{{\color{orange}\textbf{TODO:} #1 \textbf{:ODOT}\color{black}}}
\newcommand{\mysplit}{\mathsf{split}}
\newcommand{\myjoin}{\mathsf{join}}
\newcommand{\mylen}{\mathsf{len}}
\newcommand{\myread}{\mathsf{read}}
\newcommand{\mywrite}{\mathsf{write}}
\newcommand{\myset}[1]{\{#1\}}
\newcommand{\hide}[1]{ }
\newcommand{\arraylogic}{SL_{I,A,S}}

\theoremstyle{remark}
\newtheorem{definition}{Definition}[section]

@@

%% write by denghang ==================================================================
%% write by denghang ==================================================================
\hide{
    \subsection{Special CEFA for String Array}
    %%
    We define a special CEFA for string array. The CEFA uses a preserved character $``\arrconcat"$ as $array$ elements separator (note that $\arrconcat\not\in\Sigma$). The speciality of $\aut_s$ rely on  its construction.
    %%
    \paragraph{Array of fixed length and constant index} Considering an $array$ $A$ with fixed length $m$, we assume that each array element $A[i]$ $(i\in[1,m])$ is constrained by a CEFA $\aut_i\equiv (R_i, Q_i, \Sigma, \delta_i, I_i, F_i, \alpha_i)$. Note that the index of the array starts from 1.  Then we can construct the CEFA of the array $A$ as $\aut_s\equiv (R_s, Q_s, \Sigma\cup\{\arrconcat\}, \delta_s, I_s, F_s, \alpha_s)$, where:
    \begin{itemize}
        % \item $R_s = \bigcup\limits_{1\leq i \leq m} R_i$,
        \item $R_s = \bigcup\limits_{1\leq i \leq m} R_i\cup \myset{r}$ for a fresh register r,
        \item $Q_s = \bigcup\limits_{1\leq i \leq m} Q_i\cup \myset{q_0, q_f}$ for fresh states $q_0$ and $q_f$,
        \item $\delta_s$ is composed of four transitions set
              \begin{itemize}
                  \item  $\{(q_0, \arrconcat, q_1, (\myvec{0}, 1))\mid q_1\in I_1\}$,
                  \item  $\myset{(q_m, \sharp, q_f, (\myvec{0}, 0))\mid q_m\in F_m}$,
                  \item  $\{(q, \arrconcat, q', (\myvec{0}, 1)) \mid q\in F_i, q'\in I_{i+1}, i\in[1,m]\}$ where $ I_{m+1}=\emptyset$,
                  \item  and $\{(q, a, q', (\myvec{0},\myvec{v_i},\myvec{0}))\mid (q, a, q', \myvec{v_i})\in \aut_i, i\in[1,m]\}$,
              \end{itemize}
              where $(\myvec{0}, 1)$ is a vector that increases the value of $r$ by 1 and remains other registers unchanged, and $(\myvec{0},\myvec{v_i},\myvec{0})$ is a vector that updates the value of $R_i$ by $\myvec{v_i}$ and remains other registers unchanged. Similarly for $(\myvec{0}, 0)$ and so on.
        \item $I_s = \myset{q_0}$, $F_s = \myset{q_f},$ and $\alpha_s = \bigwedge\limits_{1\leq i \leq m} \alpha_i$.
              % and $\alpha_s = \bigwedge\limits_{1\leq i \leq m} \alpha_i$.
    \end{itemize}
    The resulting CEFA is shown in fig \ref{fig:fixed_len_cons_idx}, where $q_i\in I_i$ and $q_i'\in F_i$ for each $i\in [1,m]$. The main idea of the construction above is using $\arrconcat$ to separate each element of the array and one register to record the index of each element. We introduce two fresh states $q_0$ and $q_f$ to ensure each element is started and ended with at least one $\arrconcat $, which is beneficial to the construction of pre-images of string constraints (see \ref*{sec:pre_image} for more details). The accepting condition $\alpha_s$ is the conjunction of accepting conditions of each element.
    % It is obviously that the accepting word $w$ of $\aut_s$ is in language $\mathcal{L}(\aut_1)\arrconcat\mathcal{L}(\aut_2)\ ... \ \arrconcat \mathcal{L}(\aut_{m}) $.
    \begin{figure}[H]
        \centering
        \usetikzlibrary {automata, positioning, fit}
        \begin{tikzpicture}[initial text =, initial distance=3ex,
                node distance=1.2cm, auto,
                state/.style={circle, draw, inner sep=0pt, minimum size=5mm},
                accepting by double]

@@

        \end{tikzpicture}
        \caption{The CEFA of $A$ with accepting condition $\alpha_s = \bigwedge\limits_{1\leq i \leq m} \alpha_i$}
        \label{fig:fixed_len_cons_idx}
    \end{figure}
    The construction of $\aut_s$ can be used to formalize the array containing constant string. For example, for a 2-length array $A$ constrained by $``a" = \myread(A, 1)\wedge ``b" = \myread(A, 2) \wedge \mylen(A) = 2$ , where $\myread(A, m)$ read the  element of $A$ at constant index $m$ and $\mylen(A)$ is the length of $A$, we can construct the automaton of $A$ as following: \\
    $(\emptyset, \{q_0,q_1,q_1',q_2,q_2'\}, \{a,b,\arrconcat\}, \{q_0\xrightarrow[(1)]{\arrconcat}q_1, q_1\xrightarrow[(0)]{a}q_1', q_1'\xrightarrow[(1)]{\arrconcat}q_2, q_2\xrightarrow[(0)]{b}q_2'\}, \{q_0\}, \{q_2'\}, \top)$. 
    %%
    \paragraph{Array of varible length and index} \todo{TODO}
}%% write by denghang ==================================================================
%% write by denghang ==================================================================

\section{String Theory with Integer and Array $\arraylogic $}\label{sec:SL}
We defined a theory $\arraylogic$ with three sorts: \textbf{string}, \textbf{integer} and \textbf{array}. The \textit{elements} of \textbf{array} are in \textbf{string} sort, and the $index$ of \textbf{array} is in \textbf{Integer} sort. The syntax of $\arraylogic$ is defined in table \ref{tab:syntax}. A $\arraylogic$ formula $\varphi$ is a conjunction of the integer atomic formulas $t_1\circ t_2$, or the string atomic formulas $x = \myread(A,m)$, $A = \mywrite(B, x, m)$, $x=\myjoin(A,u)$, $A=\mysplit(x,e)$, $x=y\cdot z$ and $x\in e$. Moreover, $t$ is a integer term built from the constant $m$, the varible $i$, the size of array $\mylen(A)$ and the length of string $\mylen(x)$ by operators $+$ and $-$; $\circ \in \{=,\not=, <, \leq, >, \geq \}$ is the set of arithmatic binary predicates on integer terms $t_1$ and $t_2$; $e$ is a regular expression; $\myread(A, m)$ returns the $m$-th element of $A$; $\mywrite(B, x ,m)$ returns the array $B$ after $x$ is written to $m$-th position of $B$; $\myjoin(A,u)$ joins all of the elements in $A$ by string $u$; $\mysplit(x,e)$ splits $x$ into an ordered string array by regex $e$; $y\cdot z$ is the concatenation of $y$ and $z$.
%  For technical convenience, we assume that all assignments in F are \textit{Static Singly Assignment (SSA)} form.
\begin{table}
    \renewcommand{\arraystretch}{1.5}
    \begin{tabular}{l l l r}
        $x,y$ & $\cdots$ &                                                                       & String varibles          \\
        $m,n$ & $\cdots$ &                                                                       & Integer constants        \\
        $i,j$ & $\cdots$ &                                                                       & Integer varibles         \\
        $A,B$ & $\cdots$ &                                                                       & Array varibles           \\
        $t$   & $::=$    & $m \mid i \mid t+t \mid t-t \mid \mylen(A)\mid \mylen(x) $            & Integer term             \\
        $SC$  & $::=$    & $x=\myread(A, m) \mid A = \mywrite(B, x ,m) \mid x=\myjoin(A,u) \mid$ &                          \\
              &          & $A = \mysplit(x, e) \mid x= y\cdot z\mid x\in e$                      & String constraints       \\
        $\varphi$   & $::=$    & $t\circ t \mid SC \mid \varphi\wedge \varphi $                                    & Formula of $\arraylogic $

    \end{tabular}
    \caption{Syntax of $\arraylogic$}
    \label{tab:syntax}
\end{table}
\subsection{Definition of Straight-Line Formula }
\begin{definition}[Relational constraints ]
    Relational constraints $\phi$ are defined by the following rules:
    \begin{align*}
        \phi \ ::= \  & x=y\cdot z \mid x=\myjoin(A,u) \mid A = \mysplit(x,e) \mid         \\
                      & x=\myread(A, m) \mid A = \mywrite(B, x ,m) \mid \phi \wedge \phi 
    \end{align*}
\end{definition}
\begin{definition}[Straight-line relational constraints]
    A relational constraint $\phi$ is straight-line if $\phi = \bigwedge\limits_{1\leq i \leq m} \chi_i = P_i$ such that
    \begin{itemize}
        \item $\chi_1,\cdots,\chi_m $ are mutally distinct string varibles and array varibles.
        \item For each $i\in [m] $, all the varibles in $P_i$ are in the set $\{\chi_1,\cdots,\chi_{m-1} \}$.
    \end{itemize}
\end{definition}
\begin{definition}[Straight-line formula]
    A formula $F$ of $\arraylogic$ is straight-line iff all relational constraints in $F$ are straight-line.
\end{definition}

\section{Pre-images of String Constraints}\label{sec:pre_image}
In this section, we define the pre-images of string constraint formulae.
\paragraph{- $i = \mylen(A)$:} The pre-image for the array length function is a CEFA $\aut_{arrlen}=(R, Q, \Sigma\cup\{\arrconcat\}, \delta, I, F, \alpha)$, where
\begin{itemize}
    \item $R = \{r\}$ for a fresh register $r$, $Q = \{q_0, q_f\}$, $\delta = \myset{q_0\xrightarrow[(0)]{\Sigma}q_0, q_0\xrightarrow[(1)]{\sharp}q_0, q_0\xrightarrow[(0)]{\arrconcat}q_f }$, $I = \myset{q_0}$, $F = \myset{q_f}$, and $\alpha \equiv r = i$.
\end{itemize}

@@

\todo{Rewrite the pre-images of array read and write function.}
\paragraph{- $x = \myread(A, m)$:} Suppose that $x$ is constrained by the CEFA $\aut_x=(R_x, Q_x, \Sigma, \delta_x, I_x, F_x, \alpha_x)$, then the pre-image for the array read function is a CEFA $\aut_{read}=(R, Q, \Sigma\cup\{\arrconcat\}, \delta, I, F, \alpha)$, where
\begin{itemize}
    \item $R = R_x$,
    \item $Q = Q_x\cup\myset{q_0, q_1, \cdots, q_{m}}$,
    \item $\delta$ is the union of six transitions set:
          \begin{itemize}
              \item $\myset{q_i\xrightarrow[(\myvec{0})]{\Sigma}q_i\mid i\in [1,m-1]}$,
              \item $\myset{q_i\xrightarrow[(\myvec{0})]{\arrconcat}q_{i+1}\mid i\in[0,m-2]}$,
              \item $\myset{q_{m-1}\xrightarrow[(\myvec{0})]{\arrconcat}q_0^x \mid q_0^x\in I_x}$,
              \item $\myset{q_f^x\xrightarrow[(\myvec{0})]{\arrconcat}q_m \mid q_f^x\in F_x}$,
              \item $\myset{q_m\xrightarrow[(\myvec{0})]{\arrconcat}q_m}$,
              \item $\myset{q\xrightarrow[(\myvec{v})]{a}q'\mid q\xrightarrow[\myvec{v}]{a}q'\in \delta_x}$.
          \end{itemize}
    \item $I = \myset{q_0}$, $F = \myset{q_m}$, and $\alpha = \alpha_x$.
\end{itemize}

@@

\bibliographystyle{abbrv}
\bibliography{ref}
